\bf Summary

### 2025-02-11 paper/paper.qmd
use PLoS template which worx

---
title: "`traveltime`: an R package to calculate walking and motorised travel time from user-specified locations"
format:
  jasa-pdf:
    keep-tex: true   
tags: 
  - R
  - geographic information systems
  - spatial analysis
authors:
  - name: Gerard E. Ryan
    orcid: 0000-0003-0183-7630
    corresponding: true
    affiliation: "1, 2"
  - name: Nicholas Tierney
    orcid: 0000-0003-1460-8722
    affiliation: "1, 3"
  - name: Nick Golding
    orcid: 0000-0001-8916-5570
    affiliation: "1, 4"
  - name: Daniel J. Weiss
    orcid: 0000-0002-6175-5648
    affiliation: "1, 3"
affiliations:
 - name: The Kids Research Institute Australia, Nedlands 6009 WA, Australia
   index: 1
 - name: Melbourne School of Population and Global Health, University of Melbourne, 3010, VIC, Australia
   index: 2
 - name: Curtin University, Bentley, WA, Australia
   index: 3
 - name: University of Western Australia, WA, Australia
   index: 4
date: In prep 2025
bibliography: paper.bib
---

@@

[^2]: Approximately 0.008333 decimal degrees, or just below 1 km^2^ at the equator

At the other end of the scale, the calculations can require relatively large amounts of onboard memory for analyses over large landscapes (e.g. analyses over Africa required \~ 72 GB RAM). Developing methods to handle large landscapes either with less memory or via cloud resources would be helpful to make such analyses accessible to those without access to larger computing resources.

>>>>

---
title: "`traveltime`: an R package to calculate walking and motorised travel time from user-specified locations"
format:
  plos-pdf:
    keep-tex: true   
tags: 
  - R
  - geographic information systems
  - spatial analysis
authors:
  - name: Gerard E. Ryan
    orcid: 0000-0003-0183-7630
    corresponding: true
    affiliation: 1, 2
  - name: Nicholas Tierney
    orcid: 0000-0003-1460-8722
    affiliation: 1, 3
  - name: Nick Golding
    orcid: 0000-0001-8916-5570
    affiliation: 1, 4
  - name: Daniel J. Weiss
    orcid: 0000-0002-6175-5648
    affiliation: 1, 3
affiliations:
 - name: The Kids Research Institute Australia, Nedlands 6009 WA, Australia
   index: 1
 - name: Melbourne School of Population and Global Health, University of Melbourne, 3010, VIC, Australia
   index: 2
 - name: Curtin University, Bentley, WA, Australia
   index: 3
 - name: University of Western Australia, WA, Australia
   index: 4
date: In prep 2025
bibliography: paper.bib
---

@@

[^2]: Approximately 0.008333 decimal degrees, or just below 1 km$^2$ at the equator

At the other end of the scale, the calculations can require relatively large amounts of onboard memory for analyses over large landscapes (e.g. analyses over Africa required ~ 72 GB RAM). Developing methods to handle large landscapes either with less memory or via cloud resources would be helpful to make such analyses accessible to those without access to larger computing resources.

### 2025-02-11 paper/paper.tex
use PLoS template which worx

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  12pt]{article}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

@@

\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-1in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.7in}%
\addtolength{\topmargin}{-1in}%
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{agsm}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Summary},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\begin{document}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\date{January 1, 2025}
\title{\bf Summary}
\author{
Gerard E. Ryan\\
1, 2\\
and\\Nicholas Tierney\\
1, 3\\
and\\Nick Golding\\
1, 4\\
and\\Daniel J. Weiss\\
1, 3\\
}
\maketitle

\bigskip
\bigskip
\begin{abstract}

\end{abstract}

\newpage
\spacingset{1.9} % DON'T change the spacing!

@@

Statement of need

@@

Example: walking from public transport in Singapore

@@

\section{Prepare the data and friction
surface}\label{prepare-the-data-and-friction-surface}

@@

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(terra)}
\FunctionTok{library}\NormalTok{(geodata)}

\NormalTok{singapore\_shapefile }\OtherTok{\textless{}{-}} \FunctionTok{gadm}\NormalTok{(}
  \AttributeTok{country =} \StringTok{"Singapore"}\NormalTok{,}
  \AttributeTok{level =} \DecValTok{0}\NormalTok{,}
  \AttributeTok{path =} \FunctionTok{tempdir}\NormalTok{(),}
  \AttributeTok{resolution =} \DecValTok{2}
\NormalTok{)}

\NormalTok{singapore\_shapefile}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(traveltime)}
\FunctionTok{head}\NormalTok{(stations)}
\end{Highlighting}
\end{Shaded}

@@

Now that we have the two items of data that we require, the next step is
to prepare a friction surface for our area of interest. We will use the
friction surface from \citet{weiss2020} that can be downloaded by
\texttt{traveltime} with the function \texttt{get\_friction\_surface()}.
We can pass in our basemap \texttt{singapore\_shapefile}, a
\texttt{SpatVector}, directly as the \texttt{extent}. We're interested
in walking time from a station, so we'll download the walking friction
surface by specifying \texttt{surface\ =\ "walk2020"}. (Alternatively,
we could use \texttt{surface\ =\ "motor2020"} for motorised travel if
that were of interest.) We're also only interested in walking \emph{on
land} so we then mask out areas outside of the land boundary in
\texttt{singapore\_shapefile}:

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{friction\_singapore}
\end{Highlighting}
\end{Shaded}

@@

Below we plot the friction surface raster \texttt{friction\_singapore},
with the vector boundary \texttt{singapore\_shapefile} as a grey line,
and \texttt{stations} as grey points (Figure Figure~\ref{fig-data}).
\texttt{traveltime} takes resistance values of friction
\citep{gdistance2017}, so higher values of friction indicate more time
travelling across a given cell. In this case walking friction is fairly
uniform across the islands barring a few small inaccesible locations.
More heterogeneity would be expected in less developed locations with
fewer transport corridors.

@@

\section{Calculate and plot the travel
time}\label{calculate-and-plot-the-travel-time}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{trave\_time\_singapore }\OtherTok{\textless{}{-}} \FunctionTok{calculate\_travel\_time}\NormalTok{(}
  \AttributeTok{friction\_surface =}\NormalTok{ friction\_singapore,}
  \AttributeTok{points =}\NormalTok{ stations}
\NormalTok{)}

\NormalTok{trave\_time\_singapore}
\end{Highlighting}
\end{Shaded}

@@

We present the resulting calculated travel times in Figure
Figure~\ref{fig-result} where, as expected, the travel times are lowest
near station exits (per Figure Figure~\ref{fig-data}) and progressively
higher further away. Note that the results in
\texttt{trave\_time\_singapore} include infinite values (\texttt{Inf}
above). In Figure Figure~\ref{fig-data}, the islands to the south and
north-east are shown as filled cells, but unconnected with the mainland.
The raster cells for these islands appear absent in Figure
Figure~\ref{fig-result}. Because they are not connected to any cells
with a station, the calculated travel time is infinite, and so these
cells do not appear in Figure Figure~\ref{fig-result}.

@@

Opportunities for future development

@@

Firstly, \texttt{traveltime} currently has access to walking and
motorised friction surfaces for 2020, both at 30 arc-second
resolution\footnote{Approximately 0.008333 decimal degrees, or just
  below 1 km\textsuperscript{2} at the equator}. Although the user can
supply their own friction surface, we expect most applications will use
these existing surfaces given the extensive work needed in creating new
ones \citep{weiss2018global, weiss2020global}. As landscapes are not
dynamic, it may be useful to incorporate updated versions of these
friction surfaces if and when they are available, though this is likely
to occur first through \texttt{malariaAtlas}. Furthermore, although the
resolution of these data is likely to be suitable for larger landscape
foci, higher resolution data may be helpful for more locally focussed
analyses. For instance, although the example here was chosen for it's
simplicity and low computational demands, a \textasciitilde1
km\textsuperscript{2} cell size is a relatively large area to walk
across, and thus actual waking times may vary significantly within each
cell. We underline however that a user can provide their own higher
resolution friction surface at present if desired.

@@

Acknowledgements

@@

References

  \bibliography{paper.bib}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%

\PassOptionsToPackage{table}{xcolor}

\documentclass[
  10pt,
  letterpaper,
]{article}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

@@

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% marvosym package for additional characters
\usepackage{marvosym}

% cite package, to clean up citations in the main text. Do not remove.
% Using natbib instead
% \usepackage{cite}

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% create "+" rule type for thick vertical lines
\newcolumntype{+}{!{\vrule width 2pt}}

% create \thickcline for thick horizontal lines of variable length
\newlength\savedwidth
\newcommand\thickcline[1]{%
  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
  \cline{#1}%
  \noalign{\vskip\arrayrulewidth}%
  \noalign{\global\arrayrulewidth\savedwidth}%
}

% \thickhline command for thick horizontal lines that span the table
\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
\hline
\noalign{\global\arrayrulewidth\savedwidth}}

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in 
\textheight 8.75in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\renewcommand{\figurename}{Fig}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Header and Footer with logo
\usepackage{lastpage,fancyhdr}
\usepackage{epstopdf}
%\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
%\setlength{\headheight}{27.023pt}
%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\today}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[numbers,square,comma]{natbib}
\bibliographystyle{plos2015}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={traveltime: an R package to calculate walking and motorised travel time from user-specified locations},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\begin{document}
\vspace*{0.2in}

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{\texttt{traveltime}: an R package to calculate walking
and motorised travel time from user-specified
locations} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
\\
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
Gerard E. Ryan\textsuperscript{1*}, Nicholas
Tierney\textsuperscript{2}, Nick Golding\textsuperscript{3}, Daniel J.
Weiss\textsuperscript{2}
\\
\bigskip
\textbf{1} 1, 2, \\ \textbf{2} 1, 3, \\ \textbf{3} 1,
4, \\ \textbf{4} The Kids Research Institute Australia, Nedlands 6009
WA, Australia, \\ \textbf{5} Melbourne School of Population and Global
Health, University of Melbourne, 3010, VIC,
Australia, \\ \textbf{6} Curtin University, Bentley, WA,
Australia, \\ \textbf{7} University of Western Australia, WA,
Australia, 
\bigskip

% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
%
% Remove or comment out the author notes below if they aren't used.
%
% Primary Equal Contribution Note
%\Yinyang These authors contributed equally to this work.

% Additional Equal Contribution Note
% Also use this double-dagger symbol for special authorship notes, such as senior authorship.
%\ddag These authors also contributed equally to this work.

% Current address notes
%\textcurrency Current Address: Dept/Program/Center, Institution Name, City, State, Country % change symbol to "\textcurrency a" if more than one current address note
% \textcurrency b Insert second current address
% \textcurrency c Insert third current address

% Deceased author note
%\dag Deceased

% Group/Consortium Author Note
%\textpilcrow Membership list can be found in the Acknowledgments section.

% Use the asterisk to denote corresponding authorship and provide email address in note below.
* \\ email \href{mailto:gerry.ryan@thekids.org.au}{gerry.ryan@thekids.org.au}

\end{flushleft}

\linenumbers

\section{Summary}\label{summary}

@@

\section{Statement of need}\label{statement-of-need}

@@

\section{Example: walking from public transport in
Singapore}\label{example-walking-from-public-transport-in-singapore}

@@

\subsection{Prepare the data and friction
surface}\label{prepare-the-data-and-friction-surface}

@@

@@

@@

Now that we have the two items of data that we require, the next step is
to prepare a friction surface for our area of interest. We will use the
friction surface from \citet{weiss2020global} that can be downloaded by
\texttt{traveltime} with the function \texttt{get\_friction\_surface()}.
We can pass in our basemap \texttt{singapore\_shapefile}, a
\texttt{SpatVector}, directly as the \texttt{extent}. We're interested
in walking time from a station, so we'll download the walking friction
surface by specifying \texttt{surface\ =\ "walk2020"}. (Alternatively,
we could use \texttt{surface\ =\ "motor2020"} for motorised travel if
that were of interest.) We're also only interested in walking \emph{on
land} so we then mask out areas outside of the land boundary in
\texttt{singapore\_shapefile}:

@@

@@

Below we plot the friction surface raster \texttt{friction\_singapore},
with the vector boundary \texttt{singapore\_shapefile} as a grey line,
and \texttt{stations} as grey points (Figure Fig~\ref{fig-data}).
\texttt{traveltime} takes resistance values of friction
\citep{gdistance2017}, so higher values of friction indicate more time
travelling across a given cell. In this case walking friction is fairly
uniform across the islands barring a few small inaccesible locations.
More heterogeneity would be expected in less developed locations with
fewer transport corridors.

@@

\subsection{Calculate and plot the travel
time}\label{calculate-and-plot-the-travel-time}

@@

@@

We present the resulting calculated travel times in Figure
Fig~\ref{fig-result} where, as expected, the travel times are lowest
near station exits (per Figure Fig~\ref{fig-data}) and progressively
higher further away. Note that the results in
\texttt{trave\_time\_singapore} include infinite values (\texttt{Inf}
above). In Figure Fig~\ref{fig-data}, the islands to the south and
north-east are shown as filled cells, but unconnected with the mainland.
The raster cells for these islands appear absent in Figure
Fig~\ref{fig-result}. Because they are not connected to any cells with a
station, the calculated travel time is infinite, and so these cells do
not appear in Figure Fig~\ref{fig-result}.

@@

\section{Opportunities for future
development}\label{opportunities-for-future-development}

@@

Firstly, \texttt{traveltime} currently has access to walking and
motorised friction surfaces for 2020, both at 30 arc-second
resolution\footnote{Approximately 0.008333 decimal degrees, or just
  below 1 km\(^2\) at the equator}. Although the user can supply their
own friction surface, we expect most applications will use these
existing surfaces given the extensive work needed in creating new ones
\citep{weiss2018global, weiss2020global}. As landscapes are not dynamic,
it may be useful to incorporate updated versions of these friction
surfaces if and when they are available, though this is likely to occur
first through \texttt{malariaAtlas}. Furthermore, although the
resolution of these data is likely to be suitable for larger landscape
foci, higher resolution data may be helpful for more locally focussed
analyses. For instance, although the example here was chosen for it's
simplicity and low computational demands, a \textasciitilde1
km\textsuperscript{2} cell size is a relatively large area to walk
across, and thus actual waking times may vary significantly within each
cell. We underline however that a user can provide their own higher
resolution friction surface at present if desired.

@@

\section{Acknowledgements}\label{acknowledgements}

@@

\nolinenumbers
\renewcommand\refname{References}
  \bibliography{paper.bib}
